%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Medium Length Professional CV
% LaTeX Template
% Version 2.0 (8/5/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Trey Hunner (http://www.treyhunner.com/)
%
% Important note:
% This template requires the resume.cls file to be in the same directory as the
% .tex file. The resume.cls file provides the resume style used for structuring the
% document.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass{resume} % Use the custom resume.cls style

\usepackage[left=0.75in,top=0.4in,right=0.75in,bottom=0.6in]{geometry} % Document margins
\usepackage{fontawesome}
\usepackage{marvosym}

\name{R. Benjamin Upenieks} % Your name
\address
%{248 Golf Course ~$\cdot$~
 {(519)  591  9922 ~$\cdot$~   ben.upenieks@uwaterloo.ca ~$\cdot$~ \faGithub \space github.com/bupenieks ~$\cdot$~ \faExternalLink \space https://bupenieks.github.io } % Your address
%\address{} % Your secondary addess (optional)
%\address{} % Your phone number and email

\begin{document}


%----------------------------------------------------------------------------------------
%	TECHNICAL STRENGTHS SECTION
%----------------------------------------------------------------------------------------

%\begin{rSection}{Technical Strengths}

%\begin{tabular}{ @{} >{\bfseries}l @{\hspace{6ex}} l }
%Computer Languages & C, C++, Java, HTML, CSS \\ 
%Tools & git, SVN, Vim
%\end{tabular}

%\end{rSection}

%----------------------------------------------------------------------------------------
%	WORK EXPERIENCE SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Experience}
%\begin{tabular}{ @{} >{\bfseries}l @{\hspace{6ex}} l }
%Languages & C, C++, Java (Strong), JavaScript, Python (Experienced), HTML, CSS, Bash \LaTeX (familiar)
%\end{tabular}

\begin{rSubsection}{Ford Motor Company}{May -- September 2017}{Software Developer Co-op -- C, C++, Qt, QNX}{}
%\setlength\itemsep{0.1em} -- vertical space between bullets

\item Developed the backend for the in-vehicle touchscreen system with Qt. 
\item Refactored and decoupled an existing monolithic code base to provide a modular and more robust interface for vehicle services to request alerts and notifications.
\item Designed and implemented notification arbitration, queueing and suppression subsystems in highly governed environments and exposed the API to internal clients in C and C++.
\item Wrote unit tests for notification API business logic and exception flows using Google Test \& Google Mock.

\end{rSubsection}

\begin{rSubsection}{Contribution to 4th-Year Engineering Design Team}{February -- April 2017}{Diaphyseal Bone Manufacturing - C++}{}
\item Tasked with analyzing open-source C++ 3D slicing software to improve the deconstruction of SolidWorks models into functional G-Code for 3D printing
\item Proposed a solution to increase the efficiency of printing diaphyseal bone grafts on cylindrical print beds.
\end{rSubsection}

\end{rSection}


%------------------------------------------------
\begin{rSection}{Projects}


%\begin{rSubsection}{TinyHFS}{}{A UNIX-style hierarchical file storage system }{}

%\item Developed and implemented
%\end{rSubsection}


\begin{rSubsection}{TinyHFS -- Hierarchical File Storage System}{}{Arduino -- C}{github.com/Bupenieks/TinyHFS}

\item Developed and implemented a low level, navigable, bytewise hierarchical file system on an Arduino microcontroller.%with primitive UNIX-like commands
\item Complete with full CRUD and auxiliary UNIX-like operations.
%\item Interfaced with the onboard 24LC256 Microchip EEPROM via I\textsuperscript{2}C communication over the serial bus.
\end{rSubsection}


\begin{rSubsection}{BeatSync}{}{Android -- Java}{Google Play Store -- github.com/Bupenieks/BeatSync}
\item Created an app to synchronize rowing strokes with songs from your Spotify playlists using accelerometer data.
\item Includes auxiliary music player functionality with play/pause, track seeking, and playlist selection. 
\end{rSubsection}


\begin{rSubsection}{CoffeeHouse}{}{Node.js -- JavaScript -- Hack the North 2017}{github.com/Bupenieks/CoffeeHouse}
\item Writing the backend for a lightweight, Git-style version control service to enable open-source music creation. 
\end{rSubsection}


\begin{rSubsection}{Scrabble}{}{Node.js, MongoDB -- JavaScript, Python}{github.com/Bupenieks/Scrabble}
\item  Created an online two-player Scrabble clone in Node with client connection powered by MongoDB and websockets. 
\item Equipped with a ``Best Word'' algorithm written in Python to determine optimal tile placement.
	\end{rSubsection}



\end{rSection}
\begin{rSection}{Languages \& Tools}

\begin{tabular}{ @{} >{\bfseries}l @{\hspace{1ex}} l }
Strong: & \hspace{1ex} C++, Qt, Node.js, JavaScript, Java, Android SDK, Git\\
Experienced: & \hspace{1ex}  C,  Python, NumPy, Scala, MongoDB, jQuery, HTML5, CSS3, \LaTeX\\
Familiar: & \hspace{1ex}  Swift, iOS, Django, Android NDK, Vagrant, QNX, Bash, SQL
\end{tabular}


\end{rSection}




%----------------------------------------------------------------------------------------
%	EDUCATION SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Education}

{\bf University of Waterloo}, Candidate for\textbf{ \textit{Bachelor of Software Engineering}} (BSE); 2A \\
Cumulative Average : \textbf{85.17}\% -- 1A Dean's Honour List \\
Overloading courses with the intent of earning a \textit{\textbf{Joint Honours Degree in Statistics}}

\begin{tabular}{ @{} >{\bfseries}l @{\hspace{2ex}} l }
Relevant Coursework : & Coursera Deep Learning Specialization

\end{tabular}

%\begin{tabular}{ @{} >{\bfseries}l @{\hspace{1ex}} l }
%Languages: & C, C++, Java (Strong), JavaScript, Python (Experienced), HTML, CSS, Bash, \LaTeX (familiar)
%\end{tabular}

\end{rSection}

\begin{rSection}{Awards}
\begin{rSubSubsection}{}{}{}{}

\item Recognized by the 1A term's \textbf{Dean's Honour List}
%\item University of Waterloo \textbf{President's Scholarship of Distinction} -- 97\% high school average 
		%\\ -- Awarded to students admitted with an admission average of 95\% or greater
\item University of Waterloo \textbf{Engineering Entrance Scholarship} \& \textbf{President's Scholarship of Distinction} 
%\\  -- Awarded for  academic excellence and showing potential to thrive in professional skill development
\item \textbf{SAT} Subject Tests Math 1: 800 - Physics: 800 -  \textbf{Perfect scores}
\item \textbf{ACT} Math: 36 - Science: 36 - \textbf{Perfect scores}; English: 34 - Reading: 34 - Writing: 11 - 99th percentile overall
\item Trained and competed at the highest level in Canada as a soccer goalkeeper.



\item \textbf{Toronto FC} \& \textbf{Ontario Provincial Soccer Team} -- Goalkeeper for Team Ontario and Toronto FC Academy  
\end{rSubSubsection}








\end{rSection}


%\begin{rSection}{Clubs, Volunteer \& Community}
%\begin{rSubSubsection}{}{}{}{}
%\item Community Liaison for Constituency Communications; Local Member of Provincial Parliament Office 
%\item Canvassed and campaigned for the local candidate during two elections 
%\item Physics Club \& AP Calculus AB Club Founder; St. David Secondary School	
%\item Academic Tutor, Mathematics and English, grades 9 - 12				
%\item Model United Nations member; Kitchener \& SOMA participant; Toronto, Ontario\\
%\end{rSubSubsection}



%{rSection}



%----------------------------------------------------------------------------------------
%	EXAMPLE SECTION
%----------------------------------------------------------------------------------------

%\begin{rSection}{Section Name}

%Section content\ldots

%\end{rSection}

%----------------------------------------------------------------------------------------

\end{document}
